Changelog

### 2021-05-25 Chapter-1/Chapter-1.tex
Updates from Overleaf

\chapter{Introduction} \label{chapter1}
\section{Side-channels}
A side-channel attack is any attack that allows adversaries to infer sensitive information based on ``side-channel'' information. In 1996, Kocher published the first side-channel attack in the literature~\cite{kocher1996timing}. He demonstrated that an attacker could break cryptosystems like RSA or Diffie-Hellman by measuring the execution time of a simple modular exponentiation operation. The simple modular exponentiation calculates the remainder of an exponent divided by an integer.  The simple modular exponentiation operation executes faster when the exponent bit is 0 and slower when the exponent bit is 1. An attacker can infer the whole secret key by measuring the execution time of the operation. Since then, researchers have proposed numerous side-channel attacks which exploit different kinds of ``side-channel'' information, including execution timing~\cite{184415,disselkoen2017prime+}, power consumption~\cite{kar20178}, electromagnetic radiation~\cite{agrawal2002side,217605}, and even sound~\cite{chari1999towards}.

>>>>

\chapter{Introduction} \label{chapter1}
\section{Side-channels}
A side-channel attack is any attack that allows adversaries to infer sensitive information based on ``side-channel'' information. In 1996, Kocher published the first side-channel attack in the literature~\cite{kocher1996timing}. He demonstrated that an attacker could break cryptosystems such as RSA or Diffie-Hellman by measuring the execution time of a simple modular exponentiation operation. The simple modular exponentiation calculates the remainder of an exponent divided by an integer.  The simple modular exponentiation operation executes faster when the exponent bit is 0 and slower when the exponent bit is 1. An attacker can infer the whole secret key by measuring the execution time of the operation. Since then, researchers have proposed numerous side-channel attacks which exploit different kinds of ``side-channel'' information, including execution timing~\cite{184415,disselkoen2017prime+}, power consumption~\cite{kar20178}, electromagnetic radiation~\cite{agrawal2002side,217605}, and even sound~\cite{chari1999towards}.

### 2021-05-25 Chapter-3/Chapter-3.tex
Updates from Overleaf

\subsection{Performance}
The second limitation of current side-channel detection tools is the performance bottleneck, especially for those tools based on symbolic analysis. For example, CaSym can only analyze small programs. CacheD performs better than CaSym. However, it still can not handle large programs like RSA directly. It uses some domain knowledge to analyze only part of the program. The expensive cost comes from the symbolic execution. Symbolic execution can be used to explore all the possible paths, which is useful for many tasks.
Nevertheless, it is significantly less likely for cache side channel bugs in crypto primitives. Cryptography primitives are more likely to have an even coverage of different paths (based on pseudo-random cipher states), and a branch-based side channel is vulnerable from the very first branch on secret data, making complex path conditions less often relevant. Therefore, we choose to analyze a trace instead of the whole path. Another bottleneck is that the IR. While the IR can significantly decrease the difficulty of the implementation, it also gives introduce an average of 10x of the overall overhead. Third, using SMT solving can be convenient because of its generality, and it is commonly used for its precision when checking branch feasibility in symbolic execution. But the decision problems that an SMT solver handles are typically at least NP-hard. On the other hand, we can use some methods to avoid the solver and find side-channel leakages more efficiently.

@@

If the memory access equals to $t_b$, we can use the constraint $c_i :
  f_i(\beta, k, m) >> L = t_b >> L$ to model the observation on secret-dependent
data accesses. Let's take a look the examples in Figure~\ref{fig:chapter3:da}(a).
Suppose the base address of the array \textsf{T} is 10. We symbolize both the $k$ and $m$. The formula that represents the memory access at line 6 is $10 + m + (k \mod 128)$. Therefore, we have the formula: $(10 + m + (k_1 \mod 128)) >> 6  \neq (10 + m + (k_2 \mod 128)) >> 6$.
Here is one possible solution: $m = 0, k_1 = 0, k_2 = 127$. In fact, for any $m$, we can always find possible $k_1$ and $k_2$ that satisfy the above inequation. It is a true leakage. The leakage can also be identified by previous tools like CacheD. Similarly, we can model the memory access at line 6 with the formula:  $10 + m + (k_1 \mod 32) >> 6  \neq 10 + m + (k_2 \mod 32) >> 6$. If $m = 0$, we can not find satisfiable $k_1$ and $k_2$ for the above inequation, which indicates that it will always access the same cache line at line 6. However, it is possible that for some $m \in M$, we can find such a pair of $k_1$ and $k_2$ ($m = 53, k_1 = 1, k_2 = 0$). If the actual public input is $0$, existing tools (e.g., CacheD) can miss such a vulnerability. On the other hand, our tool can identify the leakages precisely.

@@

\subsubsection{Secret-dependent Control-flows}
An adversary can infer sensitive information from secret dependent control-flows. There are two kinds of control-transfer instructions: the unconditional control-transfer instructions and the conditional transfer instructions.
The unconditional instructions, like CALL, JUMP, RET transfer control from one code segment location to another. Since the transfer is independent of the input sensitive information, an attacker was not able to infer any sensitive information from the control-flow.
So the unconditional control-transfer does not leak any information based on our threat model. During the symbolic execution, we update the register information and memory cells with new formulas accordingly.

The conditional control-flow transfer instructions, like conditional jumps, depending on CPU states, may or may not transfer control flows. For conditional jumps, the CPU will test if certain condition flag
(e.g., CF = 0, ZF =1) is met and jump to certain branches, respectively.
The symbolic engine will compute the flag and represent the flag in a symbol
formula. Because we are running on a symbolic execution on an execution trace, we know which branch is executed.
If a conditional jump uses the CPU status flag, we will generate the constraint accordingly.

>>>>

\subsection{Performance}
The second limitation of current side-channel detection tools is the performance bottleneck, especially for those tools based on symbolic analysis. For example, CaSym can only analyze small programs. CacheD performs better than CaSym. However, it still can not handle large programs such as RSA directly. It uses some domain knowledge to analyze only part of the program. The expensive cost comes from the symbolic execution. Symbolic execution can be used to explore all the possible paths, which is useful for many tasks.
Nevertheless, it is significantly less likely for cache side channel bugs in crypto primitives. Cryptography primitives are more likely to have an even coverage of different paths (based on pseudo-random cipher states), and a branch-based side channel is vulnerable from the very first branch on secret data, making complex path conditions less often relevant. Therefore, we choose to analyze a trace instead of the whole path. Another bottleneck is that the IR. While the IR can significantly decrease the difficulty of the implementation, it also gives introduce an average of 10x of the overall overhead. Third, using SMT solving can be convenient because of its generality, and it is commonly used for its precision when checking branch feasibility in symbolic execution. But the decision problems that an SMT solver handles are typically at least NP-hard. On the other hand, we can use some methods to avoid the solver and find side-channel leakages more efficiently.

@@

If the memory access equals to $t_b$, we can use the constraint $c_i :
  f_i(\beta, k, m) >> L = t_b >> L$ to model the observation on secret-dependent
data accesses. Let's take a look the examples in Figure~\ref{fig:chapter3:da}(a).
Suppose the base address of the array \textsf{T} is 10. We symbolize both the $k$ and $m$. The formula that represents the memory access at line 6 is $10 + m + (k \mod 128)$. Therefore, we have the formula: $(10 + m + (k_1 \mod 128)) >> 6  \neq (10 + m + (k_2 \mod 128)) >> 6$.
Here is one possible solution: $m = 0, k_1 = 0, k_2 = 127$. In fact, for any $m$, we can always find possible $k_1$ and $k_2$ that satisfy the above inequation. It is a true leakage. The leakage can also be identified by previous tools such as CacheD. Similarly, we can model the memory access at line 6 with the formula:  $10 + m + (k_1 \mod 32) >> 6  \neq 10 + m + (k_2 \mod 32) >> 6$. If $m = 0$, we can not find satisfiable $k_1$ and $k_2$ for the above inequation, which indicates that it will always access the same cache line at line 6. However, it is possible that for some $m \in M$, we can find such a pair of $k_1$ and $k_2$ ($m = 53, k_1 = 1, k_2 = 0$). If the actual public input is $0$, existing tools (e.g., CacheD) can miss such a vulnerability. On the other hand, our tool can identify the leakages precisely.

@@

\subsubsection{Secret-dependent Control-flows}
An adversary can infer sensitive information from secret dependent control-flows. There are two kinds of control-transfer instructions: the unconditional control-transfer instructions and the conditional transfer instructions.
The unconditional instructions, such as CALL, JUMP, RET transfer control from one code segment location to another. Since the transfer is independent of the input sensitive information, an attacker was not able to infer any sensitive information from the control-flow.
So the unconditional control-transfer does not leak any information based on our threat model. During the symbolic execution, we update the register information and memory cells with new formulas accordingly.

The conditional control-flow transfer instructions, such as conditional jumps, depending on CPU states, may or may not transfer control flows. For conditional jumps, the CPU will test if certain condition flag
(e.g., CF = 0, ZF =1) is met and jump to certain branches, respectively.
The symbolic engine will compute the flag and represent the flag in a symbol
formula. Because we are running on a symbolic execution on an execution trace, we know which branch is executed.
If a conditional jump uses the CPU status flag, we will generate the constraint accordingly.

### 2021-05-25 Chapter-4/Chapter-4.tex
Updates from Overleaf

After we split those constraints into several small constraints, we count the number of solutions for each constraint. Even though the dimension has been
significantly reduced by the previous step, this is still a \#P problem. For our project, we apply the approximate counting instead of exact counting for two reasons. First, we do not need to have a very precise result of the exact number of total solutions since the information is defined with a logarithmic function. We do not need to distinguish between constraints having $10^{10}$ or $10^{10} + 10$ solutions; they are very close to after taking logarithmic. Second, the
precise model counting approaches, like Davis-Putnam-Logemann-Loveland (DPLL)
search, have difficulty scaling up to large problem sizes.

@@

Considering the cache-collision timing attack~\cite{Bonneau11894063_16}, the
probability of leakage decreases when the lookup table entry or element size gets
smaller. We analyze the table lookup with \tool{}. It is a common operation used by symmetric ciphers like AES. Figure~\ref{fig:chapter4:big_lookup} is from the original AES reference implementation. It uses a lookup table and each entry is 4 bytes, which is vulnerable to side-channel attacks. Most cryptography libraries adopted the mitigated version, shown in Figure~\ref{fig:chapter4:small_lookup}. \tool{} reports three leakage sites for
each lookup table, with 4.0 bits for the table with smaller entries and 2.0
bits for the table with bigger entries, confirming the theory.
The evaluation result shows the version with smaller lookup tables leaks less information, which confirm the effectiveness of the mitigation technique.

>>>>

After we split those constraints into several small constraints, we count the number of solutions for each constraint. Even though the dimension has been
significantly reduced by the previous step, this is still a \#P problem. For our project, we apply the approximate counting instead of exact counting for two reasons. First, we do not need to have a very precise result of the exact number of total solutions since the information is defined with a logarithmic function. We do not need to distinguish between constraints having $10^{10}$ or $10^{10} + 10$ solutions; they are very close to after taking logarithmic. Second, the
precise model counting approaches, such as Davis-Putnam-Logemann-Loveland (DPLL)
search, have difficulty scaling up to large problem sizes.

@@

Considering the cache-collision timing attack~\cite{Bonneau11894063_16}, the
probability of leakage decreases when the lookup table entry or element size gets
smaller. We analyze the table lookup with \tool{}. It is a common operation used by symmetric ciphers such as AES. Figure~\ref{fig:chapter4:big_lookup} is from the original AES reference implementation. It uses a lookup table and each entry is 4 bytes, which is vulnerable to side-channel attacks. Most cryptography libraries adopted the mitigated version, shown in Figure~\ref{fig:chapter4:small_lookup}. \tool{} reports three leakage sites for
each lookup table, with 4.0 bits for the table with smaller entries and 2.0
bits for the table with bigger entries, confirming the theory.
The evaluation result shows the version with smaller lookup tables leaks less information, which confirm the effectiveness of the mitigation technique.

### 2021-05-25 Chapter-7/Chapter-7.tex
Updates from Overleaf

\textbf{Kernel Space Side-channel Vulnerability Detection.} There are several side-channel vulnerability detection tools together with the dissertation tool. Most of the tools are designed to analyze the side-channel leakages in user-level applications like cryptographic libraries. It is reasonable since many side-channel attacks are used to break cryptographic applications. Recently, researchers also started to focus on side-channel attacks~\cite{cao2016off} inside the kernel space. However, it is non-trivial to apply existing side-channel detection tools in the Linux kernel. Take the method in the dissertation, for example. We may have the following challenges. First, our proposed tools in the dissertation rely on the dynamic binary instrumentation frameworks to collect the runtime information. As for kernel programs, a similar task can be achieved by some whole system emulators like BitBlaze~\cite{song2008bitblaze} and S2E~\cite{chipounov2012s2e}. S2E uses a modified Linux kernel and QEMU to trace the kernel components. S2E uses QEMU 1.0, which does not support the latest Linux kernel. The engineering efforts of extending S2E to work on the latest kernel are significant. Second, the kernel has very complicated control-flow graphs compared to cryptography libraries. Some vulnerabilities need to be triggered by specific input from the user space applications or devices. Our tool can not handle interruptions and exceptions. But they are common inside the kernel. On the positive side, we expect there are plenty of leakage sites inside the kernel. The dissertation provides the approaches to pick up ``severe'' leakages automatically.

>>>>

\textbf{Kernel Space Side-channel Vulnerability Detection.} There are several side-channel vulnerability detection tools together with the dissertation tool. Most of the tools are designed to analyze the side-channel leakages in user-level applications such as cryptographic libraries. It is reasonable since many side-channel attacks are used to break cryptographic applications. Recently, researchers also started to focus on side-channel attacks~\cite{cao2016off} inside the kernel space. However, it is non-trivial to apply existing side-channel detection tools in the Linux kernel. Take the method in the dissertation, for example. We may have the following challenges. First, our proposed tools in the dissertation rely on the dynamic binary instrumentation frameworks to collect the runtime information. As for kernel programs, a similar task can be achieved by some whole system emulators such as BitBlaze~\cite{song2008bitblaze} and S2E~\cite{chipounov2012s2e}. S2E uses a modified Linux kernel and QEMU to trace the kernel components. S2E uses QEMU 1.0, which does not support the latest Linux kernel. The engineering efforts of extending S2E to work on the latest kernel are significant. Second, the kernel has very complicated control-flow graphs compared to cryptography libraries. Some vulnerabilities need to be triggered by specific input from the user space applications or devices. Our tool can not handle interruptions and exceptions. But they are common inside the kernel. On the positive side, we expect there are plenty of leakage sites inside the kernel. The dissertation provides the approaches to pick up ``severe'' leakages automatically.
